---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Introduction

`linkrep` provides an easy-to-use and extensible architecture for generating linkage quality reports in R. It simplifies the process of evaluating and reporting on the quality of data linkage.  

This package aims to provide a comprehensive tool for data linkage analysts to assess the quality of their linkage processes, while also helping data providers and researchers understand linkage errors and evaluate potential biases. 

Reports generated with `linkrep` can be further customized to fit your specific needs:

- **Add new elements**: Include additional tables, figures, sections or data as required.

- **Modify content**: Customize written portions, including the Methods section, to better reflect your specific linkage processes.

- **Personalize appearance**: Adjust the report's background, layout, styles, and references to tailor the report to your needs.

# Installation

## R Studio Installation

To install `linkrep` from GitHub, begin by installing and loading the `devtools` package:
```{r install-devtools, eval = FALSE}
# install.packages("devtools")
library(devtools)
```

Afterwards, you may install the automated data linkage package using `install_github()`:
```{r install-linkage-report-package, eval = FALSE}
devtools::install_github("CHIMB/linkrep")

# You may need to install tinytex:
# install.packages("tinytex")
# tinytex::install_tinytex()
```

## Local Installation

To install `linkrep` locally from GitHub, select the most recent release from the right-hand tab on the GitHub repository page. Download the <b>Source code (zip)</b> file, then move over to RStudio. You may then run the code:
```{r local-install, eval = FALSE}
path_to_pkg <- file.choose() # Select the unmodified package you downloaded from GitHub.
devtools::install_local(path_to_pkg)

# You may need to install tinytex:
# install.packages("tinytex")
# tinytex::install_tinytex()
```

# Main Report Elements

**Summary**: Overview of the Methods section and results  

**How to Read This Report**: Provides recommendations on how to interpret the tables and figures to assess for potential biases. 

**Linkage Rate Summary**: Includes the linkage rate table which stratifies linkage rates by sociodemographic factors and other characteristics.

**Linkage Algorithm Summary**: If provided, includes tables and figures describing the linkage algorithm and its quality.

**Performance Metrics**: If provided, includes multiple performance metrics (0-100) listed in table, visualized using a radar chart.

**Background**: Describes record linkage, how it's performed, and its limitations.

**Methods**: Details the linkage process, including pre-processing and techniques used.

**Appendix**: Algorithms that were considered for testing, along with their performance metrics, can be included at the end of the report.

# Additional Information & Documentation
For detailed instructions on formatting data for the report and customizing features, refer to the [User Documentation](https://github.com/CHIMB/linkrep/blob/main/inst/docs/User_Documentation.pdf)

# Authors
- [Elizabeth Stoughton](https://github.com/stoughty111)
- [Cole Chuchmach](https://github.com/Cole-Chuchmach)
- [Barret A. Monchka](https://github.com/barretmonchka)
