---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# linkrep

`linkrep` provides an easy-to-use and extensible architecture for generating linkage quality reports in R. It simplifies the process of evaluating and reporting on the quality of data linkage.  

This package aims to provide a comprehensive tool for data linkers to assess the quality of their linkage processes, while also helping data providers and researchers understand linkage errors and evaluate potential biases. 

Reports generated with `linkrep` can be customized to fit your specific needs:

- **Add new elements**: Include additional tables, figures, sections or data as required

- **Modify content**: Customize written portions, including the Methods section, to better reflect your specific linkage processes

- **Personalize appearance**: Adjust the report's background, layout and styles to tailor the report to your needs

## Main Report Elements
**Summary**: Overview of the Methods section and results  

**How to Read This Report**: Provides recommendations on how to interpret the tables and figures to assess for potential biases  

**Linkage Rate Summary**: Includes the linkage rate table which stratifies linkage rates by sociodemographic factors and other characteristics

**Linkage Algorithm Summary**: If provided, includes tables and figures describing the linkage algorithm and its quality  

**Background**: Describes record linkage, how it's performed, and its limitations  

**Methods**: Details the linkage process, including pre-processing and techniques used

## Installation

```{r, eval=FALSE}
# install.packages("pak")
pak::pak("CHIMB/linkrep")

# You may need to install tinytex:
# install.packages("tinytex")
# tinytex::install_tinytex()
```

### Local Installation
Navigate to **Releases** on the right-hand side of the GitHub page and select the most recent release. Download the **Source code (zip)** file, then run the following code to complete installation:
```{r, eval=FALSE}
# Select the unmodified package you downloaded from GitHub:
pkg_path <- file.choose()
devtools::install_local(pkg_path)

# You may need to install tinytex:
# install.packages("tinytex")
# tinytex::install_tinytex()
```

## Additional Information
For detailed instructions on formatting data for the report and customizing features, refer to the [User Documentation](https://github.com/CHIMB/linkrep/blob/main/inst/docs/User_Documentation.pdf)

## Authors
- [Elizabeth Stoughton](https://github.com/stoughty111)
- [Barret A. Monchka](https://github.com/barretmonchka)
