---
format: 
  pdf:
    mainfont: {mainfont}
    sansfont: {sansfont}
    fontsize: {fontsize}
    fig-pos: 'H'
    fig-align: left
    toc: false
    number-sections: true
    include-in-header:
      - {background_images}
  docx:
      reference-doc: {word_template}
      toc: true
      number-sections: true
      fig-align: left
editor: source

bibliography: {references}
csl: {citation_style}

execute:
  echo: false
  
params: 
  definitions_data_path: null
  abbreviations_data_path: null
  listed_elements_generator_function_path: null
  linked_data_repr_table_path: null
  linkage_rate_table_path: null
  linkage_rates_plot_path: null
  algorithm_summary_table_path: null
  performance_measures_table_path: null
  performance_measures_plot_path: null
  performance_measures_plot_caption: null
  missingness_table_path: null
  report_title: null
  report_subtitle: null
  left_dataset_name: null
  right_dataset_name: null
  data_linker: null
  output_format: null
  project_id: nul
  num_records_left_dataset: null
  num_records_right_dataset: null
  ground_truth: null
  data_time_period: null
  num_records_linked: null
  overall_linkage_rate: null
  report_generation_date: null
  display_back_cover_page: true
  comprehensive_report: true
  linkage_package: null
  linkage_package_version: null
  linkrep_package_version: null
  R_version: null
  datastan_package_version: null

header-includes:
- \renewcommand{\thesection}{\arabic{section}.}
- \renewcommand{\thesubsection}{\arabic{section}.\arabic{subsection}}
- \usepackage{geometry}
- \geometry{paperheight=11in, paperwidth=8.5in, top = 1in, bottom = 1in, left = 1in, right = 1in}
- \usepackage{caption}
- \raggedright
- \captionsetup[figure]{justification=raggedright,singlelinecheck=false}
- \captionsetup[table]{justification=raggedright,singlelinecheck=false}
- \usepackage{etoolbox}
---

```{r load packages, include=FALSE, message=FALSE}
library(flextable)
library(xfun)
```

```{r read_data function}
#----
# read_data
#
# Performs error checking on the file path and reads in the data.
#
# @param data_path A file path to the data.
# @param parameter The name of the parameter whose path we're checking.
#
# @return The read in data.
#----
read_data <- function(data_path, parameter){
  if (!is.null(data_path)){
    if (!file.exists(data_path)){
      stop(sprintf("Invalid argument: %s. Could not find file", parameter))
    }
    if (file_ext(data_path) != "rds"){
      stop(sprintf("Invalid argument: %s. File must be an rds file", parameter))
    }
    data <- readRDS(data_path)
  } else {
    data <- NULL
  }
  return(data)
}
```

```{r read in report elements}
# if one of these is NULL it means it wasn't supplied to the function so don't output it

listed_output_function <- read_data(params$listed_elements_generator_function_path, "listed_elements_generator_function_path")

definitions_data <- read_data(params$definitions_data_path, "definitions_data_path")

abbreviations_data <- read_data(params$abbreviations_data_path, "abbreviations_data_path")

linked_data_repr_table <- read_data(params$linked_data_repr_table_path, "linked_data_repr_table_path")

linkage_rate_table <- read_data(params$linkage_rate_table_path, "linkage_rate_table_path")

linkage_rates_over_time_plot <- read_data(params$linkage_rates_plot_path, "linkage_rates_plot_path")

algorithm_summary_table <- read_data(params$algorithm_summary_table_path, "algorithm_summary_tabl_path")

performance_measures_table <- read_data(params$performance_measures_table_path, "performance_measures_table_path")

missingness_table <- read_data(params$missingness_table_path, "missingness_table_path")

if (!is.null(params$performance_measures_plot_path)){
  if (!file.exists(params$performance_measures_plot_path)){
    stop("Invalid argument: performance_measures_plot-path. Could not find file")
  }
  if (params$output_format == "pdf"){
    if (file_ext(params$performance_measures_plot_path) != "pdf"){
      stop("Invalid argument: performance_measures_plot-path. Must be a path to a PDF file")
    }
  } else {
    if (file_ext(params$performance_measures_plot_path) != "png"){
      stop("Invalid argument: performance_measures_plot-path. Must be a path to a png file")
    }
  }
}
```

```{r set_page_width function}
#----
# set_page_width
#
# Must be called in an R chunk that has results='asis' set in the header in order to work.
#
# @param orientation String indicating the desired page orientation. Allowed values are 'portrait' and 'landscape'.
# @return A prints a list of LaTeX comands.
#----
set_page_width <- function(orientation){
  if (orientation != "portrait" & orientation != "landscape"){
    stop("Invalid argument: orientation. orientation must be either 'portrait' or 'landscape'")
  }
  if (orientation == "portrait"){
    cat("\\newpage")
    cat("\\pdfpagewidth = 8.5in")
    cat("\\pdfpageheight = 11in")
    cat("\\newgeometry{top = 1in, bottom = 1in, left = 1in, right = 1in}")
  } else {
    cat("\\newpage")
    cat("\\pdfpagewidth = 11in")
    cat("\\pdfpageheight = 8.5in")
    cat("\\newgeometry{top = 1in, bottom = 3.5in, left = 1in, right = 1in}")
  }
}
```

```{r requires_landscape_page function}
#----
# requires_landscape_page
#
# Determines if the table width is too big for a portrait page
#
# @param table A flextable object
# @return A logical indicating whether the page needs to be in landscape orientation
#----
requires_landscape_page <- function(table){
  MAX_PORTRAIT_TABLE_WIDTH <- 6.5
  return (params$output_format == "pdf" & flextable_dim(table)$widths >= MAX_PORTRAIT_TABLE_WIDTH)
}
```

\newgeometry{top = 1in, bottom = 0.8in, left = 0.5in, right = 0.5in}

::: {.content-visible when-format="pdf"}
```{=tex}
\begin{titlepage}
  \begin{flushleft}
  \thispagestyle{nopagenumbers}
  \vspace*{\fill}
    \Huge
    \textbf{`r params$report_title`}\\
    \LARGE
    `r params$report_subtitle`\\
    `r if(!is.null(params$data_time_period)) params$data_time_period`\\
    \large
    \textbf{`r if(!is.null(params$project_id)) sprintf("PHRPC \\#: %s", as.character(params$project_id))`}\\
    Prepared on `r params$report_generation_date`
  \end{flushleft}
\end{titlepage}
```
\newgeometry{top = 1in, bottom = 1in, left = 1in, right = 1in}

```{=tex}
\hypersetup{linkcolor=black}
\thispagestyle{nopagenumbers}
```
:::

::: {.content-visible when-format="docx"}
{{< pagebreak >}}
:::

\setlength{\parskip}{1.5em}

This report was prepared by Elizabeth Stoughton and Barret Monchka at the George & Fay Yee Centre for Healthcare Innovation (CHI), in collaboration with the Manitoba Centre for Health Policy (MCHP) and the Provincial Information Management and Analytics (PIMA) department at Manitoba Health (MH). We gratefully acknowledge the contributions of Randy Walld of MCHP and Craig Kasper of MH for their roles in establishing the analytical and operational processes used to prepare the data for this report, and Karan Singh and Patrick Faucher of CHI Knowledge Translation for producing the graphics. Financial support was provided by CHI under the Canadian Institutes of Health Research (CIHR) Strategy for Patient Oriented Research (SPOR).  

The software used to generate this report was developed by Elizabeth Stoughton and Cole Chuchmach, under the direction of Barret Monchka, and is available as an R package (<https://github.com/CHIMB/linkrep>) under an MIT license.  

Permission is granted to adapt and reproduce the textual and graphical content of this report, in whole or in part, for non-commercial purposes related to generating data linkage quality reports, provided the original authors are appropriately attributed. If no textual or graphical content is reproduced in dynamically generated reports, then the copyright statement and suggested citation may be revised as necessary. Findings contained in this report may be reproduced provided the source is cited. All reasonable precautions have been taken by CHI to verify the accuracy of the information contained in this report, which is provided without warranty of any kind, either expressed or implied. The reader is responsible for the interpretation of the published material. In no event shall CHI, MCHP, MH, the University of Manitoba, the Government of Manitoba, or any data providers be liable for damages arising from the use of the published material or the report-generating software.  

© `r format(Sys.Date(), "%Y")` George & Fay Yee Centre for Healthcare Innovation  

\vspace*{1cm}
**Suggested Citation**

Stoughton, Elizabeth & Monchka, Barret. *`r params$report_title`: `r params$report_subtitle`*. Winnipeg, Manitoba: George & Fay Yee Centre for Healthcare Innovation; `r format(Sys.Date(), "%B %Y")`.


::: {.content-visible when-format="pdf"}
{{< pagebreak >}}

```{=tex}
\renewcommand{\contentsname}{Table of Contents}
\tableofcontents
```
:::

`r if(!is.null(definitions_data) & !is.null(listed_output_function)) "::: {.content-visible}" else "::: {.content-hidden}"`

{{< pagebreak >}}


\setlength{\parskip}{1em}

::: {.content-visible when-format="pdf"}
```{=tex}
\section*{Definition of Terms}
\addcontentsline{toc}{section}{Definition of Terms}
```

```{r definitions pdf output, results='asis'}
if (!is.null(definitions_data) & !is.null(listed_output_function) & params$output_format == "pdf"){
  listed_output_function(
    data = definitions_data, 
    output_format = params$output_format)
}
```
:::

::: {.content-visible when-format="docx"}
# Definition of Terms {.unnumbered}

```{r definitions word output, ft.align="left"}
if (!is.null(definitions_data) & !is.null(listed_output_function) & params$output_format == "docx"){
  listed_output_function(
    data = definitions_data, 
    output_format = params$output_format)
}
```
:::

:::


`r if(!is.null(abbreviations_data) & !is.null(listed_output_function)) "::: {.content-visible}" else "::: {.content-hidden}"`

{{< pagebreak >}}

::: {.content-visible when-format="pdf"}
```{=tex}
\section*{List of Abbreviations}
\addcontentsline{toc}{section}{List of Abbreviations}
```

```{r abbreviations pdf output, results='asis'}
if (!is.null(abbreviations_data) & !is.null(listed_output_function) & params$output_format == "pdf"){
  listed_output_function(
    data = abbreviations_data, 
    output_format = params$output_format)
}
```
:::

::: {.content-visible when-format="docx"}
# List of Abbreviations {.unnumbered}

```{r abbreviations word output, ft.align="left"}
if (!is.null(abbreviations_data) & !is.null(listed_output_function) & params$output_format == "docx"){
  listed_output_function(
    data = abbreviations_data, 
    output_format = params$output_format)
}
```
:::

:::

{{< pagebreak >}}

\setlength{\parskip}{1.5em}

::: {.content-visible when-format="pdf"}
```{=tex}
\section*{Summary}
\addcontentsline{toc}{section}{Summary}
```
:::

::: {.content-visible when-format="docx"}
# Summary {.unnumbered}
:::

This report details the methods used to link records from `r params$left_dataset_name` (N = `r params$num_records_left_dataset`) to those in `r params$right_dataset_name` `r ifelse(is.null(params$num_records_right_dataset), "", sprintf("(N = %s)", params$num_records_right_dataset))` and offers guidance on evaluating the quality of the linkage process. We used an iterative approach that combines deterministic and probabilistic methods to link `r params$num_records_linked` records, achieving an overall linkage rate of `r params$overall_linkage_rate`% (see @tbl-linkage_rates).  

To discuss the findings of this report, the potential impact of linkage errors on a research study, or to request guidance on adjusting for linkage errors in a data analysis, you may schedule a consultation with CHI by visiting <https://umanitoba.ca/centre-for-healthcare-innovation/> and selecting “Request a free consultation.” For additional questions regarding this data linkage, please e-mail [data.linkage@umanitoba.ca](mailto:data.linkage@umanitoba.ca).

{{< pagebreak >}}

::: {.content-visible when-format="pdf"}
```{=tex}
\section*{How to Read This Report}
\addcontentsline{toc}{section}{How to Read This Report}
```
:::

::: {.content-visible when-format="docx"}
# How to Read This Report {.unnumbered}
:::

Integrating information from multiple data sources can generate rich and comprehensive datasets that address complex research questions. However, the uncertainty inherent in linking disparate data sources---particularly when there are no unique personal identifiers in common---may introduce bias, potentially reducing the representativeness of the linked data with respect to the target study population. This potential adverse impact on study validity has led to the development of reporting guidelines for studies using linked data, which have informed the elements included in this data linkage quality report.[@bohensky2011; @gilbert2017; @pratt2020; @elstad2023; @zhao2022]  

Researchers using this linked data should familiarize themselves with the record linkage methods employed, carefully assess the representativeness of the linked sample, and consider how linkage errors may impact the interpretation of study findings. `r if (params$comprehensive_report) "The detailed information in the [Background](#sec-background) and [Methods](#sec-methods) sections will aid in interpreting the results of the linkage process" else "The detailed information in the Background and Methods sections [available online](https://umanitoba.ca/centre-for-healthcare-innovation/) will aid in interpreting the results of the linkage process"`. When disseminating research based on linked data, we encourage adherence to reporting guidelines to support critical evaluations of findings and promote research excellence.[@bohensky2011; @gilbert2017; @pratt2020]   

The [Results](#sec-results) section includes figures and tables with descriptive statistics that summarize the performance of the linkage algorithm and the quality of the linked data. To assess the suitability of the data for research purposes, consider the following recommendations:  

\setlength{\parskip}{0.5em} 
`r list_num = 1`

`r list_num`. **Examine the representativeness of the linked sample:** Use the column percentages in @tbl-linked_data_repr to compare record characteristics with their distribution in the source data. Significant differences in these distributions may affect the generalizability of study findings. 
`r list_num = list_num + 1` 


`r list_num`. **Assess potential biases in the linkage process:** Utilize the row percentages in @tbl-linkage_rates to examine the proportion of unlinked records, stratified by sociodemographic and other characteristics. For example, a significantly lower linkage rate among females compared to males may indicate issues with the linkage algorithm or data quality that could introduce selection bias into a research study. Algorithmic biases can lead to an unrepresentative sample, as reflected in @tbl-linkage_rates.
`r list_num = list_num + 1`   


`r if(!is.null(linkage_rates_over_time_plot)) "::: {.content-visible}" else "::: {.content-hidden}"`
`r list_num`. **Explore variations in linkage rates over data acquisition dates:** Refer to `r if(!is.null(linkage_rates_over_time_plot)) "@fig-linkage_rate_dist"` to identify shifts in data quality or variations in data collection practices over time. Differential linkage error with respect to acquisition dates leads to a non-random sample.
`r if(!is.null(linkage_rates_over_time_plot)) list_num = list_num + 1`  
:::  


`r if(!is.null(performance_measures_table)) "::: {.content-visible}" else "::: {.content-hidden}"`
`r list_num`. **Evaluate linkage algorithm accuracy:** Review `r if(!is.null(performance_measures_table)) "@tbl-performance_measures_tbl"` and `r if(!is.null(performance_measures_table)) "@fig-performance_measures_chart "` to assess whether the linkage algorithm performed adequately in classifying candidate record pairs as matches or non-matches. Match classification, which involves a tradeoff between match certainty (i.e., precision or PPV) and match sensitivity (i.e., recall), should be assessed in the context of your research objectives.
`r if(!is.null(performance_measures_table)) list_num = list_num + 1`
:::   


`r if(!is.null(algorithm_summary_table)) "::: {.content-visible}" else "::: {.content-hidden}"`
`r list_num`. **Critically review the multi-step linkage algorithm employed:** Analyze the parameters used in each step of the algorithm `r if(!is.null(algorithm_summary_table)) "(@tbl-algorithm_summary)"`, including the matching technique (deterministic or probabilistic), acceptance threshold, and variables considered, to evaluate the confidence that linked records belong to the same individual. For example, exact matches on multiple fields generally reduce the chances of linkage error, whereas relying heavily on approximate string matching may increase noise in the linked data. Similarly, higher acceptance thresholds and variables with strong discriminative power (e.g., last name or personal unique identifiers) tend to reduce linkage errors compared to lower classification cut-offs and variables with weaker discriminative ability (e.g., sex/gender or postal code).
:::

\setlength{\parskip}{1.5em} 
{{< pagebreak >}}

```{r linked data representativeness table page orientation before, results='asis'}
if (requires_landscape_page(linkage_rate_table)){
  set_page_width("landscape")
}
```

# Results {#sec-results}

## Linked Data Summary

```{r linked data representativeness table, ft.align="left"}
#| label: tbl-linked_data_repr
#| tbl-cap: !expr paste0("Characteristics of records in ", params$left_dataset_name, " (N = ", params$num_records_left_dataset, ifelse(params$data_time_period == "", "", ", "), params$data_time_period, ") and those that linked to ",  params$right_dataset_name, ".") 

linked_data_repr_table
```

```{r linked data representativeness table page orientation after, results='asis'}
if (requires_landscape_page(linkage_rate_table)){
  set_page_width("portrait")
}
```

{{< pagebreak >}}

```{r linkage rate table page orientation before, results='asis'}
if (requires_landscape_page(linkage_rate_table)){
  set_page_width("landscape")
}
```

## Linkage Rate Summary

```{r linkage rate table, ft.align="left"}
#| label: tbl-linkage_rates
#| tbl-cap: !expr paste0("Stratified linkage rates for records in ", params$left_dataset_name, " (N = ", params$num_records_left_dataset, ifelse(params$data_time_period == "", "", ", "), params$data_time_period, ") that linked to ",  params$right_dataset_name, ".") 

linkage_rate_table
```

```{r linkage rate table page orientation after, results='asis'}
if (requires_landscape_page(linkage_rate_table)){
  set_page_width("portrait")
}
```

`r if(!is.null(linkage_rates_over_time_plot)) "::: {.content-visible}" else "::: {.content-hidden}"`

```{r linkage rates over time plot}
#| label: fig-linkage_rate_dist
#| fig-cap: !expr paste0("Distribution of linkage rates over data acquisition dates for records in ", params$left_dataset_name,".")
#| fig-alt: A bar plot illustrating the distributions of linkage rates over acquisition dates. The y-axis represents the linkage rate percent, and the x-axis represents the acquisition dates.
if (!is.null(linkage_rates_over_time_plot)){
  linkage_rates_over_time_plot
}
```

:::

`r if (!is.null(algorithm_summary_table) | !is.null(performance_measures_table))"::: {.content-visible}" else "::: {.content-hidden}"` 

{{< pagebreak >}} 

:::

`r if(!is.null(algorithm_summary_table)) "::: {.content-visible}" else "::: {.content-hidden}"`

```{r algorithm summary table page orientation before, results='asis'}
if (!is.null(algorithm_summary_table)){
  if (requires_landscape_page(algorithm_summary_table)){
    set_page_width("landscape")
  }
}
```

## Linkage Algorithm Summary

```{r algorithm summary table, ft.align="left"}
#| label: tbl-algorithm_summary
#| tbl-cap: !expr paste0("Summary of the multi-step algorithm to link records in ", params$left_dataset_name, " to those in ", params$right_dataset_name, ".")

if (!is.null(algorithm_summary_table)){
  algorithm_summary_table
}
```

```{r algorithm summary table page orientation after, results='asis'}
if (!is.null(algorithm_summary_table)){
  if (requires_landscape_page(algorithm_summary_table)){
    set_page_width("portrait")
  }
}
```

:::

```{r performance measures page orientation before, results='asis'}
if (!is.null(performance_measures_table)){
  if (requires_landscape_page(performance_measures_table)){
    set_page_width("landscape")
  }
}
```

`r if(is.null(algorithm_summary_table) & !is.null(performance_measures_table)) "::: {.content-visible}" else "::: {.content-hidden}"`

## Linkage Algorithm Summary

:::

`r if(!is.null(performance_measures_table)) "::: {.content-visible}" else "::: {.content-hidden}"`

```{r performance measures table, ft.align="left"}
#| label: tbl-performance_measures_tbl
#| tbl-cap: !expr paste0("Classification performance for linking records in ", params$left_dataset_name, " to those in ", params$right_dataset_name, ".")

if (!is.null(performance_measures_table)){
  performance_measures_table
}
```

`r if (!is.null(params$performance_measures_plot_path)) "::: {.content-visible}" else "::: {.content-hidden}"`

![`r params$performance_measures_plot_caption`](`r params$performance_measures_plot_path`){#fig-performance_measures_chart fig-alt="A radar chart displaying the metrics listed in @tbl-performance_measures_tbl."}

:::

```{r performance measures page orientation after, results='asis'}
if (!is.null(performance_measures_table)){
  if (requires_landscape_page(performance_measures_table)){
    set_page_width("portrait")
  }
}
```

:::

`r if (!is.null(missingness_table)) "::: {.content-visible}" else "::: {.content-hidden}"` 

{{< pagebreak >}} 

:::

`r if (!is.null(missingness_table)) "::: {.content-visible}" else "::: {.content-hidden}"`

```{r missingness table page orientation before, results='asis'}
if (!is.null(missingness_table)){
  if (requires_landscape_page(missingness_table)){
    set_page_width("landscape")
  }
}
```

## Data Quality Assessment

```{r missingness table, ft.align="left"}
#| label: tbl-missing_counts
#| tbl-cap: !expr paste0("Missingness of select variables within ", params$left_dataset_name, ".")

if(!is.null(missingness_table)){
  missingness_table
}
```

```{r missingness table page orientation after, results='asis'}
if (!is.null(missingness_table)){
  if (requires_landscape_page(missingness_table)){
    set_page_width("portrait")
  }
}
```

:::

`r if(params$comprehensive_report) "::: {.content-visible}" else "::: {.content-hidden}"`

{{< include {extra_textual_content} >}}

:::

{{< pagebreak >}}

::: {.content-visible when-format="pdf"}
```{=tex}
\section*{References}
\addcontentsline{toc}{section}{References}
```
:::

::: {.content-visible when-format="docx"}
# References {.unnumbered}
:::

::: {#refs}
:::

`r if(params$display_back_cover_page) "::: {.content-visible}" else "::: {.content-hidden}"`

::: {.content-visible when-format="pdf"}
{{< pagebreak >}}

```{=tex}
\begin{titlepage}
    \begin{center}
    \thispagestyle{nopagenumbers}
        \vspace*{5cm}
        

    \end{center}
\end{titlepage}
```
:::

:::
